Predicting Success of Clinical Interviews via Probabilistic Modeling of Patient-Provider Communication Sequences

### 2017-03-09 paper/AMIA/main.tex
added subsection impact of hidden states

\subsection*{\textit{Predictive performance for alternating PPC code sequences}}
Table~\ref{tab:result_alt_seq} summarizes predictive performance of the proposed methods on alternating sequences. These results indicate that second-order HMM had better recall than MC, achieving 97.13\% recall. Similar to the case of natural sequences, first-order MC has the lowest performance compared to all other models across all metrics. Also, as in the case of natural sequences, prediction method based on second-order HMM has the highest recall of 0.9713 and the highest accuracy of 0.9713 among all models. However, different from natural sequences, second-order MC demonstrates the highest precision 0.9778 as well as the highest F-measure of 0.9736. We also observed that second-order MC and HMM outperform first-order MC and HMM. Specifically, moving from first-order to second-order MC improves precision, recall and F-measure of the MI success prediction method by 0.01\%, 2.72\% and 1.37\%, respectively. Moving from first-order to second-order HMM improves precision, recall and F-measure of the MI success prediction method by 3.32\%, 4.30\% and 4.82\%, respectively.\\

@@

 \multirow{2}{*}{Markov Chain} & First order & 0.9777 & 0.9437 & 0.9604 & 0.9437\\\cline{2-6}
 & Second order & \textbf{0.9778} & 0.9694 & \textbf{0.9736} & 0.9694\\ \hline
 \multirow{2}{*}{Hidden Markov Model} & First order & 0.9392 & 0.9313 & 0.9253 & 0.9392\\ \cline{2-6}
 & Second order & 0.9704 & \textbf{0.9713} & 0.9699  & \textbf{0.9713}\\ \hline
  \end{tabular}
\end{table}

\subsection*{\textit{Impact of hidden states}}
Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews. Table~\ref{tab:common_patterns} provides examples of typical PPC sequences that frequently appear in successful and unsuccessful motivational interviews.

\begin{table}[h]
\centering
\caption{Emission matrix}
\label{tab:emission_matrix}
  \begin{tabular}{|l|l|}
  \hline
   \textbf{Hidden} & \textbf{Log probability of each symbol (301, 117, 327, 343, 101, 305, 311, 331, 310, 328, 112,)} \\ 
   \textbf{states} & \textbf{ 326, 302, 308, 314, 306, 119, 313, 333, 307, 325, 344, 109, 106, 332, 120, 345, 309, 312, 114,} \\
    & \textbf{ 304, 115, 329, 330, 103, 116, 324, 323, 118, 303, 334)} \\\hline      
S-1 & -2.896, -4.595, -4.031, -5.584, -11.383, -2.772, -1.532, -2.78, -3.816, -2.745, -8.444, -4.42, -3.485, \\
& -2.438, -8, -5.577, -10.333, -2.216, -3.462, -3.479, -3.71, -5.615, -20.984, -7.635, -5.7, -5.694, -3.629, \\
& -3.32, -3.195, -8, -5.007, -8, -4.636, -3.67, -8, -8, -11.027, -10.448, -8, -8, -8 \\\hline
S-2 & -3.092, -1.344, -4.322, -4.707, -6.065, -2.205, -5.701, -2.064, -5.71, -3.778, -3.121, -6.97, -3.074, \\
& -3.505, -8, -3.879, -3.229, -5.707, -2.708, -4.729, -5.538, -4.993, -4.116, -3.125, -8, -3.29, -4.439, \\
& -4.414, -6.341, -8, -8, -8, -5.964, -7.009, -8, -8, -6.067, -6.071, -8, -8, -8 \\\hline
S-3 & -3.092, -1.344, -4.322, -4.707, -6.065, -2.205, -5.701, -2.064, -5.71, -3.778, -3.121, -6.97, -3.074, \\
& -3.505, -8, -3.879, -3.229, -5.707, -2.708, -4.729, -5.538, -4.993, -4.116, -3.125, -8, -3.29, -4.439, \\
& -4.414, -6.341, -8, -8, -8, -5.964, -7.009, -8, -8, -6.067, -6.071, -8, -8, -8 \\\hline
S-4 & -3.092, -1.344, -4.322, -4.707, -6.065, -2.205, -5.701, -2.064, -5.71, -3.778, -3.121, -6.97, -3.074, \\
& -3.505, -8, -3.879, -3.229, -5.707, -2.708, -4.729, -5.538, -4.993, -4.116, -3.125, -8, -3.29, -4.439, \\
& -4.414, -6.341, -8, -8, -8, -5.964, -7.009, -8, -8, -6.067, -6.071, -8, -8, -8 \\\hline

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.70\textwidth]{figures/hidden-states.png}
    \caption{\textbf{Impact of hidden states in the performance of HMM}.}
    \label{fig:hidden-states}
\end{figure}

>>>>

\subsection*{\textit{Predictive performance for alternating PPC code sequences}}
Table~\ref{tab:result_alt_seq} summarizes predictive performance of the proposed methods on alternating sequences. These results indicate that second-order HMM had better recall than MC, achieving 97.13\% recall. Similar to the case of natural sequences, first-order MC has the lowest performance compared to all other models across all metrics. Also, as in the case of natural sequences, prediction method based on second-order HMM has the highest recall of 0.9713 and the highest accuracy of 0.9713 among all models. However, different from natural sequences, second-order MC demonstrates the highest precision 0.9778 as well as the highest F-measure of 0.9736. We also observed that second-order MC and HMM outperform first-order MC and HMM. Specifically, moving from first-order to second-order MC improves precision, recall and F-measure of the MI success prediction method by 0.01\%, 2.72\% and 1.37\%, respectively. Moving from first-order to second-order HMM improves precision, recall and F-measure of the MI success prediction method by 3.32\%, 4.30\% and 3.70\%, respectively.\\

@@

 \multirow{2}{*}{Markov Chain} & First order & 0.9777 & 0.9437 & 0.9604 & 0.9437\\\cline{2-6}
 & Second order & \textbf{0.9778} & 0.9694 & \textbf{0.9736} & 0.9694\\ \hline
 \multirow{2}{*}{Hidden Markov Model} & First order & 0.9392 & 0.9313 & 0.9353 & 0.9392\\ \cline{2-6}
 & Second order & 0.9704 & \textbf{0.9713} & 0.9699  & \textbf{0.9713}\\ \hline
  \end{tabular}
\end{table}

\subsection*{\textit{Impact of hidden states}}
Since the number of hidden states may impact the classification
performance, we use an optimum number of hidden states in HMM's. Figure~\ref{fig:hidden-states} shows the performance of HMMs with a different number of hidden states. The results illustrate that, in general, the state number didn't impact the classification performance significantly. All four models show their best performance when state number in between 4-8. Table~\ref{tab:emission_matrix} shows an example of the distribution of PPC codes over the optimum number of hidden states (4) for successful (S) and unsuccessful (U) HMMs. Log probability of each code reflects the state significance. In other words, these probabilities indicate that which hidden states are close to adolescent commitment language and which are not?

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.70\textwidth]{figures/hidden-states.png}
    \caption{\textbf{Performance of HMM's with different sequence type, order, and number of hidden states used in the experiment}.}
    \label{fig:hidden-states}
\end{figure}

\begin{table}[h]
\centering
\caption{Distribution of PPC codes over the optimal hidden states for successful (S) and unsuccessful (U) HMMs.}
\label{tab:emission_matrix}
  \begin{tabular}{|l|l|}
  \hline
   \textbf{Hidden} & \textbf{Log probability of each symbol (301, 117, 327, 343, 101, 305, 311, 331, 310, 328, 112,} \\ 
   \textbf{states} & \textbf{ 326, 302, 308, 314, 306, 119, 313, 333, 307, 325, 344, 109, 106, 332, 120, 345, 309, 312, 114,} \\
    & \textbf{ 304, 115, 329, 330, 103, 116, 324, 323, 118, 303, 334)} \\\hline      
S-1 & -2.896, -4.595, -4.031, -5.584, -11.383, -2.772, -1.532, -2.78, -3.816, -2.745, -8.444, -4.42, -3.485, \\
& -2.438, -8, -5.577, -10.333, -2.216, -3.462, -3.479, -3.71, -5.615, -20.984, -7.635, -5.7, -5.694, -3.629, \\
& -3.32, -3.195, -8, -5.007, -8, -4.636, -3.67, -8, -8, -11.027, -10.448, -8, -8, -8 \\\hline
S-2 & -3.092, -1.344, -4.322, -4.707, -6.065, -2.205, -5.701, -2.064, -5.71, -3.778, -3.121, -6.97, -3.074, \\
& -3.505, -8, -3.879, -3.229, -5.707, -2.708, -4.729, -5.538, -4.993, -4.116, -3.125, -8, -3.29, -4.439, \\
& -4.414, -6.341, -8, -8, -8, -5.964, -7.009, -8, -8, -6.067, -6.071, -8, -8, -8 \\\hline
S-3 & -3.092, -1.344, -4.322, -4.707, -6.065, -2.205, -5.701, -2.064, -5.71, -3.778, -3.121, -6.97, -3.074, \\
& -3.505, -8, -3.879, -3.229, -5.707, -2.708, -4.729, -5.538, -4.993, -4.116, -3.125, -8, -3.29, -4.439, \\
& -4.414, -6.341, -8, -8, -8, -5.964, -7.009, -8, -8, -6.067, -6.071, -8, -8, -8 \\\hline
S-4 & -3.092, -1.344, -4.322, -4.707, -6.065, -2.205, -5.701, -2.064, -5.71, -3.778, -3.121, -6.97, -3.074, \\
& -3.505, -8, -3.879, -3.229, -5.707, -2.708, -4.729, -5.538, -4.993, -4.116, -3.125, -8, -3.29, -4.439, \\
& -4.414, -6.341, -8, -8, -8, -5.964, -7.009, -8, -8, -6.067, -6.071, -8, -8, -8 \\\hline

@@
